---
title: "HW05_ploter"
output: html_notebook
---

```{r}
library(tidyverse)
library(tidymodels)
library(leaps)
library(rpart)
library(rpart.plot)
```

```{r}
set.seed(39692)
```

## Problem 1
1) Divide the data into training and model comparison parts (25% for model comparison).

```{r}
df = read_delim("hw5_C39692.csv", delim = ",")

splitted=initial_split(df,prop=0.75)
train_df=training(splitted)
test_df=testing(splitted)
head(test_df)
```

2) Use the training data to find the best Decision Tree model.

```{r}
rec1=recipe(y~.,data=train_df)
tree_spec=decision_tree(mode="regression",cost_complexity=tune(),min_n=5)|>set_engine("rpart",xval=0)
wf2=workflow()|>add_recipe(rec1)|>add_model(tree_spec)
cvdata=vfold_cv(train_df,v=5)
grid=data.frame(cost_complexity=10^seq(-6,0,length.out=11))
tune_result=wf2|>tune_grid(cvdata,grid=grid,metrics=metric_set(rmse))
show_best(tune_result)
```


```{r}
mod1=rpart::rpart(y~.,data=train_df, cp=select_best(tune_result, metric="rmse")$cost_complexity,minsplit=10,method="anova")

cp_8=as_tibble(mod1$cptable) |> filter(nsplit==8)|>pull(CP)
mod1$variable.importance
```



3) Produce a graph of the 8 most important splits of the best model


```{r}
rpart.plot::rpart.plot(rpart::prune(mod1,cp=cp_8),digits=3)
```
```{r}
print(rpart::prune(mod1, cp = cp_8))
```


4) Explain by the graph, which decisions are made to and which is the final prediction (based on the graph
shown in the previous step) for observation corresponding to the row number 692 (the last three digits in your study book number) in the original data set.

```{r}
row_692 = df[692, ]
print(row_692)
```

To explain the decisions made by the decision tree model for row 692, we’ll trace through the tree based on the values in row_692:

Node 1 (root) starting from the root node:

The root node checks if X8 < 4.97.
For row 692, X8 = 1.069961, which is less than 4.97, so we go to the left subtree (Node 2).

Node 2:
This node checks if X5 < 7.15.
For row 692, X5 = 10.22865, which is greater than 7.15, so we proceed to the right subtree of this node (Node 5).

Node 5:
This node checks if X7 < 2.02.
For row 692, X7 = 4.730494, which is greater than 2.02, so we go to the right subtree of this node (Node 11).

Node 11 (Terminal Node):
We reached the leaf, so the predicted value is 4.51.

```{r}
predicted_value = predict(rpart::prune(mod1,cp=cp_8), row_692)

actual_value = row_692$y

cat("Predicted value:", predicted_value, "\n")
```


5) Compute the performance (rmse) of the best model on the comparison part.

```{r}
test_df|>mutate(.pred=predict(mod1,test_df))|>rmse(truth=y,estimate=.pred)
```

## Problem 2
1) Use the training data to find the best kNN model for predicting y by considering models with different
number of predictors. Take into account the variable importance information from the best tree model
to select predictors for the models you try, tune also with respect to the number of neighbors and with
respect to the power of the distance metric.

 X8         X5        X19         X1         X3         X7         X2         X6        X11         X9        X10 
85721.4490 57817.8329 55810.8159 53771.9689 45963.9292 31929.0868 21583.9712 21094.8735 19274.1639 13174.7422  8828.9545 
       X14        X17        X18        X13         X4        X16        X12    country 
 5884.8183  5623.6233  5577.3957  5042.4522  4449.4950  2816.6147  1894.0486   559.3667 


```{r}
# Recipe 1: Top 3 predictors
rec_knn_1 = recipe(y ~ X8 + X5 + X19, data = train_df)

# Recipe 2: Top 5 predictors
rec_knn_2 = recipe(y ~ X8 + X5 + X19 + X1 + X3, data = train_df)

# Recipe 3: Top 7 predictors
rec_knn_3 = recipe(y ~ X8 + X5 + X19 + X1 + X3 + X7 + X2, data = train_df)

# Recipe 4: Top 9 predictors
rec_knn_4 = recipe(y ~ X8 + X5 + X19 + X1 + X3 + X7 + X2 + X6 + X11, data = train_df)

recipes_list = list(rec_knn_1, rec_knn_2, rec_knn_3, rec_knn_4)

tuning_grid=expand_grid(neighbors=c(5,10,20,40),dist_power=c(0.5,1,2))

knn_spec = nearest_neighbor(
  mode = "regression",
  neighbors = tune(),
  dist_power = tune()
) |>
  set_engine("kknn")

results = lapply(recipes_list, function(rec) {
  wf = workflow() |>
    add_recipe(rec) |>
    add_model(knn_spec)
  
  wf |>
    tune_grid(
      resamples = cvdata,
      grid = tuning_grid,
      metrics = metric_set(rmse)
    )
})

best_results = lapply(results, show_best, metric="rmse")
best_results
```

2) Compute the performance of your best model for the comparison data. Which model (from the Decision
Tree model and kNN model) would you choose for predicting new data and why?



```{r}
wf_knn = workflow() |>
  add_recipe(rec_knn_3) |>
  add_model(knn_spec)

cv_results_knn =   wf_knn |>
    tune_grid(
      resamples = cvdata,
      grid = tuning_grid,
      metrics = metric_set(rmse)
    )

best_knn = cv_results_knn |>
  select_best(metric="rmse")

final_knn = finalize_workflow(wf_knn, best_knn)

fitted_knn = fit(final_knn, data = train_df)

test_df = test_df |>
  mutate(.pred = predict(fitted_knn, test_df) |> pull(.pred))

knn_rmse = rmse(test_df, truth = y, estimate = .pred)
knn_rmse
```

The kNN model has a lower RMSE (4.1044) than the Decision Tree model (5.11721). If predictive accuracy is the primary requirement, then we should choose the kNN model.

However, if results interpretability is also a requirement, we should not rule out the Decision Tree model.

