---
title: "Lab14"
output: html_notebook
---

```{r}
library(dplyr)
library(ggfortify)
library(tidyverse)
```


## Exercise 1.
Consider data set USArrests.

Look at the means and standard deviations of features of the data set USArrests. Which is the variable with highest variance?

```{r}
USArrests |> summarize(across(everything(), .fns=~mean(.x)))
USArrests |> summarize(across(everything(), .fns=~sd(.x)))
```


Find principal components by using scaling of variables. Look at the components of results. Do you understand, what is the meaning of those components?

```{r}
ex1.1 = stats::prcomp(USArrests, scale=TRUE)
glimpse(ex1.1)
```
```{r}
ex1.1$rotation
```


Verify that principal components are computed by multiplying scaled data (which can be obtained by scale(USArrests)) with matrix containing direction vectors of components). Matrix multiplication is %*% in R.

```{r}
scale(USArrests) %*% ex1.1$rotation - ex1.1$x
```


Produce a graph of of principal components, which shows state names and loading vectors

```{r}
autoplot(ex1.1, label=FALSE, loadings=TRUE, loadings.label=TRUE, x=1, y=2, scale=0, shape=TRUE)
```


Direction vectors are determined only up to multiplying with -1, so different implementations of PCA can produce different pictures. Produce a graph with first two principal components and loading vectors if all direction vectors are chosen to have opposite directions from the original result. Hint: multiply component matrix and rotation matrix with -1.

```{r}
ex1.1$x = -ex1.1$x
ex1.1$rotation = -ex1.1$rotation


autoplot(ex1.1, label=TRUE, loadings=TRUE, loadings.label=TRUE, x=1, y=2, scale=0, shape=TRUE)
```


Produce a scree plot (plot of variance of a component divided by the total variance) of PCA result

```{r}
proc_explained = ex1.1$sdev^2/sum(ex1.1$sdev^2)

ggplot(data.frame(comp_number=1:length(proc_explained), prox_explained=proc_explained)) + geom_line(aes(comp_number, proc_explained))
```


Analyze the results if scaling of variables were not used.

```{r}
ex1.2 = stats::prcomp(USArrests, scale=FALSE)
glimpse(ex1.2)
```
```{r}
ex1.2$rotation
```



## Exercise 2.
Consider data set ISLR::NCI60.

use glimpse to see the format of the data set

```{r}

glimpse(ISLR::NCI60)
data2 = ISLR::NCI60$data
```


```{r}
ISLR::NCI60 |> summarize(across(everything(), .fns=~mean(.x)))
ISLR::NCI60 |> summarize(across(everything(), .fns=~sd(.x)))
```

Look at the standard deviations of numerical features. what are the smallest and largest values of standard deviations of gene expressions?

```{r}
as.data.frame(data2) |> summarize(across(everything(), .fns=~mean(.x)))

sd_vals = apply(data2, 2, sd) #as.data.frame(data2, 2, sd) |> summarize(across(everything(), .fns=~sd(.x)))
range(sd_vals)

summary(sd_vals)
```


Find principal components without scaling variables. How many genes have loadings in the first component which are larger than 0.06 in absolute value?
Produce a pairwise graphs of first three principal components which show the different types of cancers with different colors. How well one can distinguish different cancer types by two-dimensional projections?

```{r}
ex2.1 = stats::prcomp(data2)
v1 = ex2.1$rotation[,1]

sum(abs(v1) > 0.06)
```

```{r}
range(v1)
```

Produce scree plot explained variance plot for the PCA method. How many components are needed to explain more than 80% of variance in the data?
Repeat the previous steps by applying PCA to normalized data. Which approach gives better graphs?

```{r}
autoplot(ex2.1, data=as.tibble(data2) |> mutate(labels=ISLR::NCI60$labs), label=FALSE, loadings=FALSE, x=1, y=2, colour='labels')
```


## Exercise 3 (for extra practice)
Install package rsvd and load the dataset digits from it. Note that in the first column there are labels of actual handwritten digits, which we leave out of PCA; remaining columns are pixels of digits.

Look at the first digit by forming a 28*28 matrix of the pixels and then by using the command image(digit_matrix[,28:1], col = gray(255:0 / 255))
Check if on the pairwise plots of the first 3 principal components it is possible to separate different digits (select 50 digits from each class for producing graphs)