---
title: "Lab16"
output: html_notebook
---

```{r}
library(dplyr)
library(ggfortify)
library(tidyverse)
library(tidymodels)

library(factoextra)
library(cluster)
library(mclust)
```


```{r}
set.seed(20231221)
n=100
x=1:n/20
y=exp(-3+x+(x-3)^2+rnorm(n,sd=0.1))
train_df=tibble(x=x,C=sample(c("a","b","c",NA),n, replace=TRUE,prob=c(0.4,0.3,0.2,0.1)),y=y)
new_data=tibble(x=c(0,1),C=c(NA,"b"))
correct_predictions=-3+new_data$x+(new_data$x-3)^2
```




## Exercise 1
Perform the transformations manually. Make sure that transformations of new data work correctly and that predictions are accurate


```{r}
table(train_df$C)
```

```{r}
get_mode = function (x) {
  tmp = names(sort(table(x)))[1]
}

transformed_df = train_df |> mutate(x=x-mean(x), C=coalesce(C, get_mode(C)), x2=x^2, y=log(y))

model=lm(y~., data=transformed_df)

# INCORRECT
#transformed_new = new_data |> mutate(x=x-mean(x), C=coalesce(C, get_mode(C)), x2=x^2) #, y=log(y))
# BETTER use constants from train dataset
transformed_new = new_data |> mutate(x=x-mean(train_df$x), C=coalesce(C, get_mode(train_df$C)), x2=x^2) #, y=log(y))

# Prediction is incorrect because every constant should be taken from train sampe (mode).
rbind(predict(model, newdata=transformed_new), correct_predictions)
```

Correct approach

## Exercise 2
Make a recipe for required transformations. Perform the task by using this recipe for data transformation before model fitting and before predictions



```{r}
# mutate compute mean of the new data
# another edge case when transform to factor variable
rec_transf = recipe(y~., data=train_df) |> 
  step_log(y, skip=TRUE) |>
  step_center(x) |>
  step_mutate(x2=x^2) |>
  step_impute_mode(C)

rec_prepared=rec_transf|> prep(train_df)
transformed_data = rec_prepared|>bake(NULL)
model = lm(y~., data=transformed_data)

transformed_new = bake(rec_prepared, new_data)
transformed_new

rbind(predict(model, newdata=transformed_new), correct_predictions)
```


## Exercise 3
Include the transformations in a workflow, fit a model for training data and make predictions for new data.

```{r}
# mutate compute mean of the new data
# another edge case when transform to factor variable
rec_transf = recipe(y~., data=train_df) |> 
  step_log(y, skip=TRUE) |>
  step_center(x) |>
  step_mutate(x2=x^2) |>
  step_impute_mode(C)

wf=workflow()|> add_recipe(rec_transf) |> add_model(linear_reg())
wf_trained = wf|> fit(train_df)

wf_trained|>predict(new_data)
```
