---
title: "Homework 01"
output: html_notebook
---

```{r}
library(tidyverse)
library(GGally)
```


## Exercise 1

Check visually the file for presence of missing values. Use read_delim command (indicating the symbol for missing values if needed) to read in the data set Carseats_mod2.csv into a variable Carseats (the command should be included in your solution).

```{r}
Carseats <- read_delim("Carseats_mod2.csv", delim = ";", na = ".", skip = 3)
```

```{r}
Carseats
```


* Make sure that all variables that, according to description, should be numeric are numeric after reading data in.

```{r}
Carseats <- Carseats |>
  mutate(Sales = as.numeric(Sales),
         CompPrice = as.numeric(CompPrice),
         Income = as.numeric(Income),
         Advertising = as.numeric(Advertising),
         Population = as.numeric(Population),
         Price = as.numeric(Price),
         Age = as.numeric(Age),
         Education = as.numeric(Education))
```



```{r}
Carseats
```

## Exercise 2

Use summary() function to look at summary information about variables so that you see also list of values of character columns. 

```{r}
summary(Carseats|>mutate(across(where(is.character),factor)))
```


* Do you see anything which may indicate data errors? What? How many missing values are in the data set? (Hint - summary function works well with factor variables)

ShelveLoc variable has categories `Bad`, `Good`, and `Medium`. However, there is an entry labeled lowercase "bad", which is likely a data entry error. This should be corrected as it represents the same category as `Bad`.

Age Max. value is 126.0 highly unlikely.

Missing Values:

* ShelveLoc variable has 1 missing value.
* CompPrice variable has 2 missing values.
* Income variable has 1 missing value.
* Price variable has 2 missing values.

Total 6 missing values across 4 columns.

## Exercise 3

Use na.omit() command to define a new data set Carseats2 where rows with missing values in the original data are left out

```{r}
Carseats2 <- na.omit(Carseats)
```

```{r}
summary(Carseats2|>mutate(across(where(is.character),factor)))
```

## Exercise 4

Use select() command to select columns from Sales to Advertising and from Price to Age from Carseats2 and apply ggpairs() function to the the result. 

```{r}
selected_data <- Carseats2 |> select(Sales:Advertising, Price:Age)
```

```{r}
ggpairs(selected_data, progress=FALSE)
```

* Are any possible data errors visible from the output?

Column ShelveLoc has 4th level category "bad" which should be replaced with "Bad".

## Exercise 5

Use mutate() command with if_else() to correct data error in one of the factor variables, store the corrected data with the same name Carseats2

```{r}
Carseats2 <- Carseats2 |> mutate(ShelveLoc = if_else(ShelveLoc == "bad", "Bad", ShelveLoc))
```

```{r}
summary(Carseats2|>mutate(across(where(is.character),factor)))
```
## Exercise 6

* Use summarize() command from dplyr package to compute a table with mean price for each combination of Urban and ShelveLoc

```{r}
mean_price_summary <- Carseats2 |>
  group_by(Urban, ShelveLoc) |>
  summarize(mean_price = mean(Price), .groups='drop')
```

```{r}
print(mean_price_summary)
```

## Exercise 7

Use ggplot() with geom_point to produce a scatter plot of Age (on x-axis) and Sales, color points according to ComPrice. Discuss the graph - is it sowing expected relationships between variables?

```{r}
ggplot(Carseats2, aes(x = Age, y = Sales, color = CompPrice)) +
  geom_point() +
  labs(title = "Scatter Plot of Age vs Sales",
       x = "Age",
       y = "Sales",
       color = "CompPrice")
```

Plot does not show any obvious relationships between Age and Sales or between Sales and CompPrice.
We can observe an outlier with an age of more than 125, which might be an error in the data.
